Modeling Time-Variable Elemental Abundances in Coronal Loop Simulations

### 2024-06-27 src/tex/ms.tex
[showyourwork] overleaf sync

In this work, we have not attempted to include the changes in abundances due to ponderomotive forces of Alfv\'en waves propagating from the corona to chromosphere, which is thought to be the cause of the FIP effect \citep{laming2015}, where low FIP elements like iron are generally enhanced in coronal loops.  Observations of such waves have been measured (\textit{e.g.} \citealt{murabito2024}), and shown to be consistent with producing the FIP effect.  However, it is not clear what timescales are required to produce the fractionation of elements, as the modeling work to date has been time-independent.  Assuming that the time-scales are short, then this is an important effect that should additionally be incorporated to properly model time-variable abundances.  One additional complexity is that for large enough Poynting fluxes, the resistive dissipation of Alfv\'en waves in the chromosphere would also cause significant amounts of evaporation \citep{reep2016}.  Estimates of Poynting fluxes suggest that waves might carry up to $\approx 10^{10}$ erg s$^{-1}$ cm$^{-2}$ in flares \citep{russell2023}, which is large enough to drive explosive evaporation and produce photospheric abundances in general, but this is an upper limit.   The net effect on elemental abundances is not clear, therefore, and more work is required to determine the interplay.

>>>>

In this work, we have not attempted to include the changes in abundances due to ponderomotive forces of Alfv\'en waves propagating from the corona to chromosphere, which is thought to be the cause of the FIP effect \citep{laming2015}, where low FIP elements like iron are generally enhanced in coronal loops.  Observations of such waves have been measured (\textit{e.g.} \citealt{murabito2024}), and shown to be consistent with producing the FIP effect.  However, it is not clear what timescales are required to produce the fractionation of elements, as the modeling work to date has been time-independent.  Assuming that the time-scales are short, then this is an important effect that should additionally be incorporated to properly model time-variable abundances.  One additional complexity is that for large enough Poynting fluxes, the resistive dissipation of Alfv\'en waves in the chromosphere would also cause significant amounts of evaporation \citep{reep2016}.  Estimates of Poynting fluxes suggest that waves might carry up to $\approx 10^{10}$ erg s$^{-1}$ cm$^{-2}$ in flares \citep{russell2023}, which is large enough to drive explosive evaporation and produce photospheric abundances in general, but this is an upper limit.   The net effect on elemental abundances is not clear, therefore, and more work is required to determine the interplay.  \textbf{Additionally, examination of stellar flares on stars with FIP-biased coronae (like the Sun) compared to inverse FIP-biased coronae (like HR 1099, \citealt{brinkman2001}) suggests that chromospheric evaporation is consistent with the change in abundances during the flares relative to quiescent times \citep{nordon2008}.  Further modeling work should also, therefore, examine how abundances change during stellar flares.}
